Slitasje og kjørespor

### 2022-12-14 DRAFT_slitasjeOgKjørespor.Rmd
text

@@

I want to look at if the indicator values can be averages over smaller
and more meaningful areas (ecological homogeneous areas) than the 5
default regions.

```{r}
path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp",
               "/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp")
```

```{r}
soner <- sf::read_sf(path)
```

```{r}
tm_shape(soner)+
  tm_polygons(col = "NAVN",
              title = "Vegetation zone")
```

It could make sense to combine some zones, but I will later split the
zones by region, and then merge sones that are very uncommon in a
spesific region.

@@

I can scale the indicator for each polygon, or I can chose to aggregate
them first. If the scaled value is representative and precise at the
polygon level, then I could scale at that level. However, the combined
surveyed area is still so little of the total area of Norway, so that
only producing indicator values for these leaves the indicator without
much value for regional assessments. Later when we produce assessments
and indicator values at the regional level, we cannot simply do an area
weighting of the polygons in each region. This is because we don't want
to assume that the polygons are representative outside of the mapped
area.

@@

-   each *dekningskart* (map of surveyed area)
-   each vegetation zone (e.g. boreonemal zone in Østlandet; excluding
    zone with too little data points)
-   each vegetation region (aka bioregion, e.g. boreonemal zone and
    continental section in Østlandet)
-   each pixel of a give size, say 1 x 1 km

The last option is problematic because we often will have too few points
to be able to estimate credible mean. I can try the second option first,
and see how many data points we get for each zone.

>>>>

#### Infrastructure index

@@

I want to look at wheter the indicator values can be averages over
smaller and more meaningful areas (ecological homogeneous areas) than
the 5 default regions. One option is this data set with vegetation zone.

```{r, eval=F}
path <- ifelse(dir == "C:", 
               "R:/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp",
               "/data/R/GeoSpatialData/BiogeographicalRegions/Norway_VegetationZones_Moen/Original/Vector/soner.shp")
```

```{r, eval=F}
soner <- sf::read_sf(path)
```

```{r, eval=F}
tm_shape(soner)+
  tm_polygons(col = "NAVN",
              title = "Vegetation zone")
```

It could make sense to combine some zones, but I will later split the
zones by region, and then merge zones that are very uncommon in a
specific region.

@@

I can scale the indicator for each polygon, or I can chose to aggregate
them first. If the scaled value is representative and precise at the
polygon level, then I could scale at that level. However, the combined
surveyed area is still so little of the total area of Norway, so that
only producing indicator values for these leaves the indicator without
much value for regional assessments. Later when we produce assessments
and indicator values at the regional level, we cannot simply do an area
weighting of the polygons in each region. This is because we don't want
to assume that the polygons are representative far outside of the mapped
area. But perhaps we can assume them to be representative inside
*homogenous ecological areas*.

@@

1.  each *dekningskart* (map of surveyed area)

2.  each vegetation zone (e.g. boreonemal zone in Østlandet; excluding
    zones with too little data points)

3.  each vegetation region (aka bioregion, e.g. boreonemal zone and
    continental section in Østlandet)

4.  each region cross tabulated agains infrastructure index classes

5.  each pixel of a give size, say 1 x 1 km

Option 5 is problematic because we often will have too few points to be
able to estimate credible mean. The second and third option can be good
for certin biotic variables that depends on local species pools ect.
However, slitasje is perhaps best explained by human presence, and
option 4 therefore sounds reasonable.
